### 2018-01-29 submission/manuscript.Rmd
incorporate nick suggestions - abstract

# What the Taq? The Influence of Different Hi-Fidelity Taq Polymerase on 16S rRNA Gene Sequencing

@@

**Background.** Increasing research has found that various methodological steps can have an impact on the observed microbial community when using 16S rRNA gene surveys. These components include, but are not necessarily limited to, preservation media, extraction kit, bead beating time, and primers. Both cycle number and Hifidelity (HiFi) DNA polymerase are sometimes overlooked during this conversation of bias considerations. Here we critically examine both cycle number and HiFi DNA polymerase for biases that may influence downstream diversity measures of 16S rRNA gene surveys.

**Methods.** DNA from Fecal samples (n = 4) were extracted using a single PowerMag DNA extraction kit with a 10 minute bead beating step and amplified at 15x, 20x, 25x, 30x, and 35x using Accuprime, Kappa, Phusion, Platinum, or Q5 HiFi DNA polymerase. Mock communities (technical replicates n = 4) consisting of the whole genome of 8 different bacteria were also amplified using the same approach. The number of OTUs (Operational Taxonomic Units) was first examined for both fecal samples and mock communites. Next, Bray-Curtis index, the error rate, sequence error prevalence, and chimera prevalence, and the chimera prevalence correlation with number of OTUs was assessed for either cycle number or HiFi DNA polymerase dependent differences.

@@

**Conclusions.** Due to the impact of HiFi DNA polymerase on the number of OTUs, common diversity metrics that incorporate this value could give artificially inflated numbers due to higher undetected chimeras. So, when designing 16S rRNA gene survey studies it is important to consider both the cycle number and the type of HiFi DNA polymerase that will be used since it can increase or decrease the number of OTUS that are observed.

>>>>

# What the Taq? The Influence of High Fidelity DNA Polymerase on 16S rRNA Gene Sequencing

@@

**Background.** Increasing research has found that various methodological steps can have an impact on the observed microbial community when using 16S rRNA gene surveys. These components include, but are not necessarily limited to, preservation media, extraction kit, bead beating time, and primers. Both cycle number and high fidelity (HiFi) DNA polymerase are sometimes overlooked when sources of bias are being considered. Here we critically examine both cycle number and HiFi DNA polymerase for biases that may influence downstream diversity measures of 16S rRNA gene surveys.

**Methods.** DNA from Fecal samples (n = 4) were extracted using a single PowerMag DNA extraction kit with a 10 minute bead beating step and amplified at 15x, 20x, 25x, 30x, and 35x using Accuprime, Kappa, Phusion, Platinum, or Q5 HiFi DNA polymerase. Mock communities (technical replicates n = 4) consisting of previously isolated whole genomes of 8 different bacteria were also amplified using the same PCR amplification approach. First, the number of OTUs (Operational Taxonomic Units) was examined for both fecal samples and mock communites. Next, Bray-Curtis index, the error rate, sequence error prevalence, and chimera prevalence were assessed based on cycle number and HiFi DNA polymerase. Finally, the chimera prevalence correlation with number of OTUs was assessed for both cycle number and HiFi DNA polymerase dependent differences.

@@

**Conclusions.** Due to the impact of HiFi DNA polymerase on the number of OTUs, common diversity metrics that incorporate this value could give artificially inflated numbers due to higher undetected chimeras. So, when designing 16S rRNA gene survey studies it is important to consider both the cycle number and the type of HiFi DNA polymerase that will be used since it can increase or decrease the number of OTUs that are observed.

### 2018-01-29 submission/manuscript.tex
incorporate nick suggestions - abstract

\section{What the Taq? The Influence of Different Hi-Fidelity Taq
Polymerase on 16S rRNA Gene
Sequencing}\label{what-the-taq-the-influence-of-different-hi-fidelity-taq-polymerase-on-16s-rrna-gene-sequencing}

@@

\textbf{Background.} Increasing research has found that various
methodological steps can have an impact on the observed microbial
community when using 16S rRNA gene surveys. These components include,
but are not necessarily limited to, preservation media, extraction kit,
bead beating time, and primers. Both cycle number and Hifidelity (HiFi)
DNA polymerase are sometimes overlooked during this conversation of bias
considerations. Here we critically examine both cycle number and HiFi
DNA polymerase for biases that may influence downstream diversity
measures of 16S rRNA gene surveys.

\textbf{Methods.} DNA from Fecal samples (n = 4) were extracted using a
single PowerMag DNA extraction kit with a 10 minute bead beating step
and amplified at 15x, 20x, 25x, 30x, and 35x using Accuprime, Kappa,
Phusion, Platinum, or Q5 HiFi DNA polymerase. Mock communities
(technical replicates n = 4) consisting of the whole genome of 8
different bacteria were also amplified using the same approach. The
number of OTUs (Operational Taxonomic Units) was first examined for both
fecal samples and mock communites. Next, Bray-Curtis index, the error
rate, sequence error prevalence, and chimera prevalence, and the chimera
prevalence correlation with number of OTUs was assessed for either cycle
number or HiFi DNA polymerase dependent differences.

@@

\textbf{Conclusions.} Due to the impact of HiFi DNA polymerase on the
number of OTUs, common diversity metrics that incorporate this value
could give artificially inflated numbers due to higher undetected
chimeras. So, when designing 16S rRNA gene survey studies it is
important to consider both the cycle number and the type of HiFi DNA
polymerase that will be used since it can increase or decrease the
number of OTUS that are observed.

>>>>

\section{What the Taq? The Influence of High Fidelity DNA Polymerase on
16S rRNA Gene
Sequencing}\label{what-the-taq-the-influence-of-high-fidelity-dna-polymerase-on-16s-rrna-gene-sequencing}

@@

\textbf{Background.} Increasing research has found that various
methodological steps can have an impact on the observed microbial
community when using 16S rRNA gene surveys. These components include,
but are not necessarily limited to, preservation media, extraction kit,
bead beating time, and primers. Both cycle number and high fidelity
(HiFi) DNA polymerase are sometimes overlooked when sources of bias are
being considered. Here we critically examine both cycle number and HiFi
DNA polymerase for biases that may influence downstream diversity
measures of 16S rRNA gene surveys.

\textbf{Methods.} DNA from Fecal samples (n = 4) were extracted using a
single PowerMag DNA extraction kit with a 10 minute bead beating step
and amplified at 15x, 20x, 25x, 30x, and 35x using Accuprime, Kappa,
Phusion, Platinum, or Q5 HiFi DNA polymerase. Mock communities
(technical replicates n = 4) consisting of previously isolated whole
genomes of 8 different bacteria were also amplified using the same PCR
amplification approach. First, the number of OTUs (Operational Taxonomic
Units) was examined for both fecal samples and mock communites. Next,
Bray-Curtis index, the error rate, sequence error prevalence, and
chimera prevalence were assessed based on cycle number and HiFi DNA
polymerase. Finally, the chimera prevalence correlation with number of
OTUs was assessed for both cycle number and HiFi DNA polymerase
dependent differences.

@@

\textbf{Conclusions.} Due to the impact of HiFi DNA polymerase on the
number of OTUs, common diversity metrics that incorporate this value
could give artificially inflated numbers due to higher undetected
chimeras. So, when designing 16S rRNA gene survey studies it is
important to consider both the cycle number and the type of HiFi DNA
polymerase that will be used since it can increase or decrease the
number of OTUs that are observed.
